---
title: "Staphylococcus aureus NuoL gene, rawdata - Aureuome Methodology"
output: html_document
---

**Loading packages** 
```{r}
library(Rcpp)
library(dada2)
library(phyloseq)
library(ggplot2)
library(Biostrings)
library(microbiome)
library(devtools)
library(BiocGenerics)
library(plyr)
library(tidyverse)
library(dplyr)
library(cowplot)
library(xlsx)
library(readxl)
library(reshape2) 
library(dplyr)
library(ggthemes)
library(RColorBrewer)
```


**Set color theme**
```{r}
theme_set(theme_bw())

color_seqCountable<-c("#AEB6BF","#E32828","#D6EAF8")

color_seqCountable_negativeControls<-c("#AEB6BF","#1B4F72","#17202A")


my.col_others <- c("CC22"="#bcbcbc",
"CC15"="#f4d6be",
"M2"="#d9d2e9",
"CC80"="#f4cccc",
"M1"="#ff835c",
"M3"="#FFEBEE",
"M5"="#DCEDC8",
"M7"="#deeaee",
"M8"="#fbcadb")

my.col_Mixed <- c("CC15/CC97"="#bdcebe","M2/CC1/M1/-"="#e6e2d3","CC22/CC8/-"="#87bdd8","M2/CC1"="#36486b","M2/CC80"="#7888e0")

my.col_ASV <- c("ASV_1" = "#c2d4dd","ASV_2"="#8ca3a3","ASV_3"="#bdcebe","ASV_4"="#ada397","ASV_5"="#b9936c","ASV_6"="#dac292","ASV_7"="#e6e2d3","ASV_8"="#87bdd8","ASV_9"="#92a8d1","ASV_10"="#fff2cc","ASV_11"="#ebab79","ASV_12"="#d5f4e6","ASV_13"="#bdcebe","ASV_14"="#ff9696","ASV_15"="#c2d4dd","ASV_16"="#36486b","ASV_17"="#ada397","ASV_18"="#ada397","ASV_19"="#b9936c","ASV_20"="#7888e0","ASV_21" = "#c2d4dd","ASV_22"="#82E0AA","ASV_23"="#bdcebe","ASV_24"= "#838060","ASV_25"="#8ca3a3","ASV_26" = "#c2d4dd","ASV_27"="#ada397","ASV_28"="#fff2cc","ASV_29"="#ffef96","ASV_30"="#36486b","ASV_31" = "#c2d4dd")

my.col_CC <- c("CC1" = "#c2d4dd","CC30/-"="#8ca3a3","CC15/CC97"="#bdcebe","CC5"="#ada397","CC398"="#b9936c","CC45"="#dac292","M2/CC1/M1/-"="#e6e2d3","CC22/CC8/-"="#87bdd8","M6"="#92a8d1","CC8"="#fff2cc","CC97"="#ebab79","CC59/-"="#d5f4e6","CC15/CC97"="#bdcebe","M4"="#ff9696","CC1"="#c2d4dd","M2/CC1"="#36486b","CC5"="#ada397","CC5"="#ada397","CC398"="#b9936c","M2/CC80"="#7888e0","CC1" = "#c2d4dd","CC121/-"="#82E0AA","CC15/CC97"="#bdcebe","-"= "#838060","CC30/-"="#8ca3a3","CC1" = "#c2d4dd","CC5"="#ada397","CC8"="#fff2cc","-"="#ffef96","M2/CC1"="#36486b","CC1" = "#c2d4dd")

my.col_ST<- c("ST1/ST188/ST3926/ST81/ST474" = "#c2d4dd","ST30/ST36/ST34/ST243/ST130/ST39/ST700/ST7317-"="#8ca3a3","ST15/ST582/ST1535/ST2126/ST1153/ST3249"="#bdcebe","ST5/ST105/ST764/ST228/ST125"="#ada397","ST398/ST1232"="#b9936c","ST45/ST508/ST3154"="#dac292","ST772/ST20/ST12/ST672/ST1281/ST1156/ST573/ST901/ST1460"="#e6e2d3","ST22/ST72/ST7"="#87bdd8","ST291/ST3126/ST5561/ST6272"="#92a8d1","ST8/ST239/ST609/ST612/ST250/ST247"="#fff2cc","ST97/ST352/ST464/ST71/ST1687/ST2459/ST1179/ST2219"="#ebab79","ST59/ST338/ST87/ST4513"="#d5f4e6","ST15/ST582/ST1535/ST2126/ST1153/ST3249"="#bdcebe","ST509/ST7750/ST89"="#ff9696","ST1/ST188/ST3926/ST81/ST474"="#c2d4dd","ST25/ST1292/ST26/ST2971/ST3011/ST834"="#36486b","ST5/ST105/ST764/ST228/ST125"="#ada397","ST5/ST105/ST764/ST228/ST125"="#ada397","ST398/ST1232"="#b9936c","ST6/ST1217/ST1613/ST6266"="#7888e0","ST1/ST188/ST3926/ST81/ST474" = "#c2d4dd","ST121/ST151/ST504/ST1210/ST3140/ST705/ST51"="#82E0AA","ST15/ST582/ST1535/ST2126/ST1153/ST3249"="#bdcebe","ST7754"= "#838060","ST30/ST36/ST34/ST243/ST130/ST39/ST700/ST7317"="#8ca3a3","ST1/ST188/ST3926/ST81/ST474" = "#c2d4dd","ST5"="#ada397","ST1073"="#fff2cc","ST1073"="#ffef96","ST25/ST1292/ST26/ST2971/ST3011/ST834"="#36486b","ST126/ST7693" = "#c2d4dd")
  
my.col_CC_from_culture <- c("CC1"="#c2d4dd","CC12"="#e6e2d3","CC15"="#bdcebe","CC182"="#DCEDC8","CC22"="#87bdd8","CC30"="#8ca3a3","CC398"="#b9936c","CC45"="#dac292","CC5"="#ada397","CC509"="#ff9696","CC59"="#d5f4e6","CC7"="#87bdd8","CC8"="#fff2cc","CC97"="#ebab79","Culture negative"="#C39BD3","NA"="#000000","Unknown"="#EE524E")

Color_samplesite2<-c("#AEB6BF","#3498DB","#85C1E9","#1F618D","#58D68D","#E32828")
```


**Load data**
```{r}
taxtable=read.table("/Volumes/data/MPV/projects/microbiome/Aureome_development/DADA2_pipelines/NuoL1_blast_output/NuoL1_seq74_tax_and_counts_BLAST_detailed copy.csv",header = TRUE,sep = ",") # NB, der er manuelt blevet rettet i navnet for AD51_SYST_v1_LS_CMO_2019_n_S84, da der ved en fejl var indskrevet et n2 istedet for n.

new_taxtable = data.frame()
for (i in 1:length(taxtable$blast_assignment)) {
  new_taxtable = rbind(new_taxtable,strsplit(taxtable$blast_assignment[i],";")[[1]])
}
colnames(new_taxtable) = c("Species","Clonal_complex","Sequence_Type","Sequence")
rownames(new_taxtable) = taxtable$ASV
new_asvtable= taxtable[,-c(1:4)]
rownames(new_asvtable) = taxtable$ASV
```

**Create new dataset**
```{r}
meta_smed<-read_excel("/Volumes/data/MPV/projects/microbiome/Aureome_development/Lab/Primary samples/Meta_data_from_Sofie.xlsx")
meta_spa<-read_excel("/Volumes/data/MPV/projects/microbiome/Aureome_development/Lab/Primary samples/Samples_Seq74_KBJA.xlsx")
qPCRtable<-read_excel("/Volumes/data/MPV/projects/microbiome/Aureome_development/data/Rnotebooks_KBJA/qPCR/aPCRtable_all.xlsx")

qpcr<-qPCRtable %>% filter(!is.na(SampleID)) %>% select(SampleID,Copy_number) %>%
  distinct(SampleID,.keep_all = TRUE) %>%
  filter(SampleID!="Neg_lys") %>%
  mutate(Log_copynumber=log10(Copy_number))


##Samlpe_data, separation based on sampleID
sam_df = data.frame(SampleID = colnames(new_asvtable)) %>% 
  mutate(Type = case_when(grepl("Mock",SampleID) ~ "Mock",
                        grepl("Neg",SampleID) ~ "Neg control",
                        grepl("NEG",SampleID) ~ "Neg control",
                        grepl("_LS",SampleID) ~ "LS skin",
                        grepl("_NLS_",SampleID) ~ "NLS skin",
                        grepl("Nose",SampleID) ~ "Nose",
                         grepl("NOSE",SampleID) ~ "Nose",
                         grepl("nose",SampleID) ~ "Nose",
                          TRUE ~ "Sample"))

sam_df$SampleID_v2<-gsub(pattern="nose",replacement = "Nose",sam_df$SampleID)
sam_df$SampleID_v2<-gsub(pattern="NOSE",replacement = "Nose",sam_df$SampleID_v2)

sam_df<-sam_df %>%
  separate(SampleID_v2,c("SampleID_short","remove"),sep="_n",remove=FALSE) %>%
  select(-remove) %>%
  left_join(meta_smed,by=c("SampleID_short"="sample_id")) %>%
  left_join(meta_spa,by="SampleID_short") %>%
  left_join(qpcr,by=c("SampleID_short"="SampleID")) %>%
  mutate(replicate=case_when(
    grepl("_n_",SampleID) ~"sample",
    grepl("_n1_",SampleID) ~"sample",
    grepl("_n2_",SampleID) ~"duplicate",
  ))


sam_df<-sam_df %>% mutate(Mock=case_when(
  SampleID_short=="MockE1" ~"Even mock",
  SampleID_short=="MockE2" ~"Even mock",
  SampleID_short=="MockE3" ~"Even mock",
  SampleID_short=="MockE4" ~"Staggered mock",
 ))

sam_df<-sam_df %>% mutate(ADnose_Healthynose=case_when(
  samplesite=="AD_LS" ~"AD_LS",
  samplesite=="AD_NLS" ~"AD_NLS",
  samplesite=="AD_nose" ~"AD_Nose",
  samplesite=="AD_Nose" ~"AD_Nose",
  samplesite=="AD_NOSE" ~"AD_Nose",
  samplesite=="Control_nose" ~"Healthy_Nose",
  Type=="Neg control" ~"Negative control",
  Type=="Mock" ~"Mock"))

sam_df<-sam_df %>% mutate(seqc=case_when(
  samplesite=="AD_LS" ~"Primary samples",
  samplesite=="AD_NLS" ~"Primary samples",
  samplesite=="AD_nose" ~"Primary samples",
  samplesite=="AD_Nose" ~"Primary samples",
  samplesite=="AD_NOSE" ~"Primary samples",
  samplesite=="Control_nose" ~"Primary samples",
  Type=="Neg control" ~"Negative control",
  Type=="Mock" ~"Mock"))

rownames(sam_df) = sam_df$SampleID
```


**Create phyloseq**
```{r}
ps<-phyloseq::phyloseq(otu_table(new_asvtable,taxa_are_rows = TRUE),sample_data(sam_df),tax_table(as.matrix(new_taxtable)))
sample_data(ps)$SeqCount<-sample_sums(ps)
```

**Sample counts**
```{r}
##Read amount in each sample
sort(sample_sums(ps))
```


**Read counts among samples and controls**
```{r}
library(ggplot2)
library(cowplot)

color_seqCountable<-c("#58D68D","#E32828","#D6EAF8")

df <- data.frame(sample_data(ps)) 
df$LibrarySize <- sample_sums(ps)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))   

df1<-ggplot(data=df, aes(x=Index, y=LibrarySize,color=seqc)) + 
  geom_point(alpha =2, size=2.5)+
  theme_bw()+
  theme(legend.position="none")+
  labs(y="Read counts",x="SampleID",color="seqc") + 
  ggtitle("Read Counts")+
  theme(axis.text.x=element_text(angle=0, size=8)) +scale_fill_manual(values = color_seqCountable)+theme(
  panel.background = element_rect(fill = "white",
                                colour = "white")) + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))+scale_colour_manual(values = color_seqCountable, limits = c("Mock", "Negative control", "Primary samples"))

df2 <- ggplot(data=df, aes(x=Index, y=LibrarySize,color=seqc)) + 
  geom_point(alpha =2, size=2.5)+
  theme_bw()+theme(legend.position="right")+
  labs(y="Read counts",x="SampleID",color="seqc") + 
  ggtitle("Read Counts zoomed in")+
  theme(axis.text.x=element_text(angle=0, size=8)) +scale_fill_manual(values = color_seqCountable)+theme(
  panel.background = element_rect(fill = "white",
                                colour = "white")) + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))+
  ylim(0,1000)+geom_hline(yintercept = 500,color="black",size=0.5)+scale_colour_manual(values = color_seqCountable, limits = c("Mock", "Negative control", "Primary samples"))+
  xlim(0, 80)

Fig_SeqCount<-plot_grid(df1,df2,labels=c('A', 'B'),rel_heights = c(30,1))

Fig_SeqCount

ggsave(Fig_SeqCount, device = "png", width = 20, height = 15, units = "cm", file ="/Users/#82E0AA bjergandersen/Desktop/Fig_SeqCount.png")
```
**Median seq_counts within sample types**
```{r}
df %>% group_by(ADnose_Healthynose) %>%
  summarise(median=median(SeqCount),min=min(SeqCount),max=max(SeqCount))
```
**All primary samples combined**
```{r}
df %>% filter(group %in% c("AD","Control")) %>% summarise(median=median(SeqCount),min=min(SeqCount),max=max(SeqCount))
```
**Test of negative controls**
```{r}
df<-data.frame(sample_data(ps))

SuppFig_Neg<-ggplot(df %>% filter(Type=="Neg control"), aes(x=SampleID, y=SeqCount, color=SampleID_short)) + 
  geom_point(alpha =2, size=2.5)+
  ylim(0,1000)+ggtitle("Negative controls")+
  theme(axis.text.x=element_text(angle=90, size=8))+theme(
  panel.background = element_rect(fill = "white",
                                colour = "white")) + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))+labs(fill = "negative control")+scale_colour_manual(values = color_seqCountable_negativeControls, limits = c("Eswab_NEG", "Neg_lys", "Neg_MM"))

SuppFig_Neg

#ggsave(SuppFig_Neg, device = "png", width = 20, height = 15, units = "cm", file ="/Users/kirstinebjergandersen/Desktop/SuppFig_Neg.png")

```

**Relative abundance**
```{r}
relative<-transform_sample_counts(ps,function(x) x/sum(x))
ps_df<-psmelt(relative) %>% arrange(OTU,Sample)
ps_df_rel<-psmelt(relative) %>% arrange(OTU,Sample) %>% mutate(Abundance=Abundance*100)

#ps_df_rel %>% arrange(Sample)
```

**total number of included primary samples, including duplicates**
```{r}
tmp<-ps_df_rel %>% filter(group %in% c("AD","Control")) %>% distinct(Sample,.keep_all = TRUE) %>% mutate(samplesite=ifelse(samplesite=="AD_NOSE","AD_nose",samplesite))
length(unique(tmp$Sample))
```

**Total number of samples, not including the duplicate samples**
```{r}
tmp2<- tmp %>% filter(replicate=="sample")
length(unique(tmp2$Sample))
```

```{r}
table(tmp2$samplesite) 
```

**Total number of samples included after removing all samples with less than 500 counts (NB, dupliates inluded here)**
```{r}
tmp3<-tmp %>% filter(SeqCount>500)
length(unique(tmp3$Sample))
#172-139. # 33 samples below 500...
# 33/172*100 #19%
```

**How many unique primary smaples inluded after removing the low-count samples**
```{r}
tmp4 <- tmp3 %>% arrange(desc(SeqCount)) %>% distinct(SampleID_short,.keep_all = TRUE)

length(tmp4$Sample)
table(tmp4$samplesite)
```

**How many unique patients included**
```{r}
length(unique(tmp4 %>% filter(group=="AD") %>% pull(id)))
```

**Read counts correlation with saample sites**
```{r}
dfny<-df %>% mutate(samplesite2=case_when(
  samplesite=="AD_LS" ~"AD_LS",
  samplesite=="AD_NLS" ~"AD_NLS",
  samplesite=="AD_nose" ~"AD_Nose",
  samplesite=="AD_Nose" ~"AD_Nose",
  samplesite=="AD_NOSE" ~"AD_Nose",
  samplesite=="Control_nose" ~"Healthy_Nose",
  Type=="Neg control" ~"Negative control",
  Type=="Mock" ~"Mock"))
  
  
p<-ggplot(dfny, aes(x=samplesite2,y=SeqCount,color=samplesite2))+
  geom_boxplot()+
  geom_jitter()+
  theme(axis.text.x=element_text(angle=0, size=8))+labs(y="Read counts",x="SampleID")+theme(
  panel.background = element_rect(fill = "white",
                                colour = "white")) + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))+scale_colour_manual(values = Color_samplesite2, limits = c("AD_LS", "AD_NLS", "AD_Nose", "Healthy_Nose","Mock","Negative control"))
Fig_seqCount_Samplesite<-p

Fig_seqCount_Samplesite
#
ggsave(Fig_seqCount_Samplesite, device = "png", width = 20, height = 15, units = "cm", file ="/Users/#82E0AA bjergandersen/Desktop/Fig_seqCount_Samplesite.png")

```
**Mocks**
```{r}
##Seq73
##Load data_seq73_nuoL
taxtable73N = read.table("/Volumes/data/MPV/projects/microbiome/Aureome_development/DADA2_pipelines/NuoL1_blast_output/NuoL1_seq73Rerun_tax_and_counts_BLAST_detailed_max_target_seqs1.csv",header = TRUE,sep = ",")

#new_taxtable = as.matrix(lapply(strsplit(taxtable$blast_assignment)))

new_taxtable73N = data.frame()
for (i in 1:length(taxtable73N$blast_assignment)) {
  new_taxtable73N = rbind(new_taxtable73N,strsplit(taxtable73N$blast_assignment[i],";")[[1]])
}

colnames(new_taxtable73N) = c("Species","Clonal_complex","Sequence_Type","Sequence")
rownames(new_taxtable73N) = taxtable73N$ASV
new_asvtable73N= taxtable73N[,-c(1:5)]
rownames(new_asvtable73N) = taxtable73N$ASV
colnames(new_asvtable73N) = c("ID2_Nose","ID3_AD_LS","ID3_AD_N","ID4_AD_LS","ID4_AD_N","NC","Mock_1A","Mock_1B","Mock_1C","Mock_2A","Mock_2B","Mock_2C","Mock_3A","Mock_3B","Mock_3C","ID1_Nose")

##Create new dataset_seq73_nuoL
sam_df73N = data.frame(SampleID = colnames(new_asvtable73N)) %>% 
  mutate(Type = case_when(grepl("Mock",SampleID) ~ "Mock",
                          SampleID == "NC" ~ "Neg control",
                          TRUE ~ "Sample"))
rownames(sam_df73N) = sam_df73N$SampleID

##Create phyloseq_seq73_nuoL
ps73N<-phyloseq::phyloseq(otu_table(new_asvtable73N,taxa_are_rows = TRUE),sample_data(sam_df73N),tax_table(as.matrix(new_taxtable73N)))


##Relative abundance_seq73_nuoL
relative73N<-transform_sample_counts(ps73N,function(x) x/sum(x))
ps_df73N<-psmelt(relative73N) %>% arrange(OTU,Sample)
ps_df_rel73N<-psmelt(relative73N) %>% arrange(OTU,Sample) %>% mutate(Abundance=Abundance*100)
```

**Taxaplot**
```{r,fig.width=15}
p1<-ggplot(ps_df_rel73N%>% filter(Type=="Mock") %>% filter(Abundance>0.5))+aes(x=Sample, y=Abundance, fill=Clonal_complex)+
  geom_bar(aes(),stat="identity",position="stack")+
  theme_bw()+
  theme(legend.position="none")+
  ggtitle("g998 test run")+
  theme(axis.text.x=element_text(angle=90, size=8))+
  labs(y="Relative abundance (%)",fill="Lineage")+
  theme(legend.text = element_text(size=10))+theme(strip.background =element_rect(fill="white"))+ scale_fill_manual(values=my.col_CC)+ theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))

p2<-ggplot(ps_df_rel%>% filter(Type=="Mock") %>% filter(Abundance>0.5))+aes(x=Sample, y=Abundance, fill=Clonal_complex)+
  geom_bar(aes(),stat="identity",position="stack")+
  theme_bw()+
  theme(legend.position="right")+
  ggtitle("g998 second run")+
  theme(axis.text.x=element_text(angle=90, size=8))+
  labs(y="Relative abundance (%)",fill="Lineage")+
  theme(legend.text = element_text(size=10))+theme(strip.background =element_rect(fill="white"))+ scale_fill_manual(values=my.col_CC)+ theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))

fig_mock <-plot_grid(p1, p2, labels=c('A', 'B'),rel_heights = c(10,15),align = 'h')

fig_mock

ggsave(fig_mock, device = "png", width = 30, height = 15, units = "cm",file="/Users/kirstinebjergandersen/Desktop/fig_mock.png")
```

**qPCR**
```{r}
dfny2<-dfny %>% mutate(Seqthreshold=ifelse(SeqCount<500,"Excluded","Included")) %>% filter(!samplesite2 %in% c("Mock","Negative control"))
cor.test(x=log10(dfny2$SeqCount),y=dfny2$Log_copynumber,method="spearman")
```

**Correlation between read counts and absolute abundance**
```{r}
p<-ggplot(dfny2 %>% filter(replicate=="sample"), aes(x=log10(SeqCount),y=Log_copynumber))+
geom_point()+
  geom_smooth(method="lm")+
  geom_vline(xintercept =log10(500),color="black")+
  theme(axis.text.x=element_text(angle=0, size=8))+labs(y="Absolute abundance (Log10)",x="Read count (log10)")+theme(
  panel.background = element_rect(fill = "white",
                                colour = "white")) + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))+scale_colour_manual(values = Color_samplesite2, limits = c("AD_LS", "AD_NLS", "AD_Nose", "Healthy_Nose"))
p

ggsave(p, device = "png", width = 20, height = 15, units = "cm", file ="/Users/kirstinebjergandersen/Desktop/qPCR.png")

```
**Statistical analysis of correlation between seqcount and copy number removing replicates**
```{r}
dfny2_sub<-dfny %>% filter(replicate=="sample")
cor.test(x=log10(dfny2_sub$SeqCount),y=dfny2_sub$Log_copynumber,method="spearman")
```

**Comparison of absolute abundance and readcounts in included/excluded samples**
```{r}
color_qPCR<-c("#9FA8DA","#2ECC71")

p<-ggplot(dfny2, aes(x=Seqthreshold,y=Log_copynumber,color=Seqthreshold))+
  geom_boxplot()+
  geom_jitter()+
  theme(axis.text.x=element_text(angle=0, size=8))+labs(y="Absolute abundance (Log10)",x="")+theme(
  panel.background = element_rect(fill = "white",
                                colour = "white")) + theme(panel.border = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black")) + scale_colour_manual(values = color_qPCR)
p

ggsave(p, device = "png", width = 20, height = 15, units = "cm", file ="/Users/kirstinebjergandersen/Desktop/qPCP.png")

```


```{r}
dfny %>% arrange(SampleID_short) %>% select(SampleID,Copy_number)
```

**Statistical test of included/excluded samples**
```{r}
wilcox.test(x=dfny2$Log_copynumber[dfny2$Seqthreshold=="Included"],y=dfny2$Log_copynumber[dfny2$Seqthreshold=="Excluded"],paried=FALSE)
```

```{r}
dfny2 %>% group_by(Seqthreshold) %>%
  summarise(median=median(Copy_number),min=min(Copy_number),max=max(Copy_number))
```

**Barplot showing CC types as well as seqdepth - With low abundant lineages (without 8% thresholds)**
```{r,fig.width=20,fig.height=15}
library(ggplot2)

meta_out<-data.frame(sample_data(ps))
meta_out<-meta_out[with(meta_out,order(SeqCount)),]
order<-meta_out$SampleID
ps_df_rel$SampleID<-factor(ps_df_rel$SampleID,levels=order)
meta_out$SampleID<-factor(meta_out$SampleID,levels=order)

meta_out$inclusion_threshold<-ifelse(meta_out$SeqCount<500,"excluded","included")
ps_df_rel$inclusion_threshold<-ifelse(ps_df_rel$SeqCount<500,"excluded","included")

##Seqcount
p1<-ggplot(meta_out%>% filter(Type!="Neg control")%>% filter(Type!="Mock") %>%arrange(SeqCount))+aes(x=SampleID, y=log10(SeqCount))+
  geom_bar(aes(),stat="identity",position="stack")+
  theme_bw()+
  theme(axis.text.x=element_blank())+
  #theme(axis.text.x=element_text(angle=90, size=8))
  labs(y="Sequence counts")+
  #geom_hline(yintercept = 500,color="red")+
  #geom_vline(xintercept = "AD39_NLS_n2_S98",color="red",size=1)+
  facet_grid(.~inclusion_threshold,scales="free_x",space="free")+
  theme(strip.background =element_rect(fill="white"))+
  theme(legend.position="none")+ 
  theme(axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.text.y = element_blank())+ theme(legend.text = element_text(size=15))

##CC plot
p2<-ggplot(ps_df_rel%>% filter(Type!="Neg control") %>% filter(Type!="Mock")%>% arrange(SeqCount))+aes(x=SampleID, y=Abundance, fill=Clonal_complex)+
  geom_bar(aes(),stat="identity",position="stack")+
  theme_bw()+
  theme(legend.position="bottom")+
  theme(axis.text.x = element_blank())+
  scale_fill_manual(values=my.col_CC)+
  labs(y="Relative abundance (%) of S. aureus",fill="S. aureus CC lineages")+
  theme(legend.text = element_text(size=15))+theme(strip.background =element_rect(fill="white"))+
  facet_grid(.~inclusion_threshold,scales="free_x",space="free")+ 
  theme(axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.text.y = element_blank())+ theme(legend.text = element_text(size=15))

# lower bar
sub<-ps_df_rel[,c("SampleID","inclusion_threshold","Type","CC_from_culture")]
sub_m <- melt(sub, id.vars = c("SampleID","Type","inclusion_threshold")) 
sub_m$SampleID<-factor(sub_m$SampleID,levels=order)


p3<-ggplot(sub_m %>% filter(Type!="Neg control") %>% filter(Type!="Mock"), aes(x=SampleID, y=variable))+
  geom_tile(aes(fill=value),width = 1.0)+ 
  theme(axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.text.y = element_blank())+ theme(legend.text = element_text(size=15))+
  scale_fill_manual(values = c(my.col_CC_from_culture))+
  labs(y="Spa-typed CC")+
  facet_grid(.~inclusion_threshold,scales="free_x",space="free")+
  theme(strip.background =element_rect(fill="white"))+
  theme(legend.position="bottom")

seqCount_CC_SPA<-cowplot::plot_grid(p1,p2,p3,align = 'v',ncol=1,rel_heights = c(3,20,6),axis = 'l')
seqCount_CC_SPA

ggsave(seqCount_CC_SPA, device = "png", width = 60, height = 25, units = "cm", file ="/Users/kirstinebjergandersen/Desktop/seqCount_CC_SPA_rawdata.png")
```

**Test of duplictes, PCoA of paired samples**
```{r}
#make new column with mutate
library(ggrepel)
dfP <- df %>% group_by(SampleID_short) %>% mutate(paired = n()) %>% mutate(SampleID2 = SampleID) %>% tibble::column_to_rownames("SampleID2")

sample_data(ps) <- sample_data(dfP)

ps_paired <- subset_samples(ps,  paired==2 & !Type %in% c("Neg control", "Mock"))
```

```{r,fig.height=3,fig.width=3}
ps_paired_hell <- transform_sample_counts(ps_paired, function(x) sqrt(x / sum(x))) #hellinger transform

ps_paired_hell_ord <- ordinate(ps_paired_hell, method = "PCoA", distance = "bray", na.rm=T)

ord_plot <- plot_ordination(ps_paired_hell, ps_paired_hell_ord, type="samples", color="SampleID_short", axes = 1:2)

coul <- c(brewer.pal(12, "Set3"), brewer.pal(12,"Paired"))
coul <- colorRampPalette(coul)(length(unique(sample_data(ps_paired_hell)$SampleID_short)))


set.seed(1)

beta <-ord_plot  + geom_point(size=3) +
  theme(legend.position = "none") +
  facet_wrap(.~ADnose_Healthynose, ncol=2) +
  geom_line(aes(group=SampleID_short)) +
  scale_color_manual(values=coul) + 
  theme(panel.background = element_rect(fill = "white",
                                colour = "white")) + 
  theme(  panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          axis.line = element_line(size = 0.5, linetype = "solid",
                                   colour = "black"))
  #geom_text_repel(aes(label=SampleID_short), max.overlaps = Inf, size = 4, force=1) 

#ggsave(beta, device = "png", width = 40, height = 30, units = "cm", file ="/Users/#82E0AA bjergandersen/Desktop/dobbeltbestemmelser_PCoA.png")

```

**Test of duplicates, CC lineafge barplot and read counts, sorted based on sampleID**
```{r,fig.height=8,fig.width=15}
meta_out<-data.frame(sample_data(ps_paired))
meta_out<-meta_out[with(meta_out,order(SampleID_short)),]
order<-meta_out$SampleID

relative_paired<-transform_sample_counts(ps_paired,function(x) x/sum(x))
ps_df_rel_paired<-psmelt(relative_paired) %>% arrange(OTU,Sample) %>% mutate(Abundance=Abundance*100)

ps_df_rel_paired$SampleID<-factor(ps_df_rel_paired$SampleID,levels=order)
meta_out$SampleID<-factor(meta_out$SampleID,levels=order)

p1<-ggplot(meta_out)+aes(x=SampleID, y=log10(SeqCount))+
  geom_bar(aes(),stat="identity",position="stack")+
  theme_bw()+
  theme(axis.text.x=element_blank())+
  labs(y="Sequence counts")+geom_hline(yintercept = log10(500),color="black",size=1)+
  #geom_vline(xintercept = "AD39_NLS_n2_S98",color="red",size=1)+
  #facet_grid(.~inclusion_threshold,scales="free_x",space="free")+
  theme(strip.background =element_rect(fill="white"))+
  theme(legend.position="none")+ 
  theme(axis.ticks.x=element_blank(),
        axis.ticks.y=element_blank(),
        axis.title.x = element_blank(),
        axis.text.x=element_blank(),
        axis.text.y = element_blank())+ theme(legend.text = element_text(size=15))

p2<-ggplot(ps_df_rel_paired)+aes(x=Sample, y=Abundance, fill=Clonal_complex)+
  geom_bar(aes(),stat="identity",position="stack")+
  theme_bw()+
  theme(legend.position="bottom")+
  #ggtitle("Clonal complexes")+
  theme(axis.text.x=element_text(angle=90, size=8))+
  labs(y="Relative abundance (%)",fill="Lineage")+
  #theme(legend.position = "right", legend.text = element_text(size=12))+
  #guides(fill = guide_legend(ncol = 1))+
  theme(legend.text = element_text(size=15))+theme(strip.background =element_rect(fill="White"))+scale_fill_manual(values=my.col_CC)

  

Paired_seqCount_barplot<-cowplot::plot_grid(p1,p2,align = 'v',ncol=1,rel_heights = c(1,7))
Paired_seqCount_barplot

ggsave(Paired_seqCount_barplot, device = "png", width = 60, height = 25, units = "cm", file ="/Users/kirstinebjergandersen/Desktop/Paired_seqCount_barplot.png")


```

**Bray-Curtis dissimilarities of paired and non-paired samples as boxplots**
```{r}
bray_dist_paired <- phyloseq::distance(ps_paired_hell, method = "bray")

bray_mat_paired <- as.matrix(bray_dist_paired)
meltBR_paired<-melt(bray_mat_paired)

#Matrix with paired sample comparisons
Paired_MAT <- meltBR_paired %>% filter(grepl("_n_",Var1)) %>% filter(grepl("_n2_", Var2)) %>% left_join(dfP %>% mutate(Var1=SampleID, Var1_SampleID_short = SampleID_short) %>% select(Var1, Var1_SampleID_short)) %>% 
left_join(dfP %>% mutate(Var2=SampleID, Var2_SampleID_short = SampleID_short) %>% select(Var2, Var2_SampleID_short)) %>% 
filter(Var1_SampleID_short == Var2_SampleID_short) %>% 
select(-Var1_SampleID_short, -Var2_SampleID_short) %>% 
mutate(name = "Paired samples")

#hist(Paired_MAT$value)
```

```{r}
#without paired samples
ps_unpaired <- subset_samples(ps, !Type %in% c("Neg control", "Mock") & sample_sums(ps) !=0) #there is one sample with 0 counts (AD27_LS_n_S150) that is removed

ps_hell_unpaired <- transform_sample_counts(ps_unpaired, function(x) sqrt(x / sum(x))) #hellinger transform

bray_dist_unpaired <- phyloseq::distance(ps_hell_unpaired, method = "bray")

bray_mat_unpaired <- as.matrix(bray_dist_unpaired)
meltBR_unpaired<-melt(bray_mat_unpaired)

#Matrix with paired sample comparisons
Unpaired_MAT <- meltBR_unpaired %>% 
left_join(dfP %>% mutate(Var1=SampleID, Var1_SampleID_short = SampleID_short) %>% select(Var1, Var1_SampleID_short)) %>% 
left_join(dfP %>% mutate(Var2=SampleID, Var2_SampleID_short = SampleID_short) %>% select(Var2, Var2_SampleID_short)) %>% 
filter(Var1_SampleID_short != Var2_SampleID_short) %>% 
select(-Var1_SampleID_short, -Var2_SampleID_short) %>% 
mutate(name = "Unpaired samples")

```

**Boxplot illustrating paired and unpaired samples**
```{r}
boxplotting<-bind_rows(Paired_MAT, Unpaired_MAT)
#Defining the order, can change if needed
boxplotting$name<-factor(boxplotting$name, levels= c("Paired samples", "Unpaired samples"))

boxp <- ggplot(boxplotting, aes(x=name, y=value)) +  
#geom_jitter(size=3, width = 0.1, alpha = 0.2) + 
  geom_boxplot(outlier.colour = NA, alpha=0.5, lwd=1) + 
  labs(y="Bray-Curtis dissimilarities \n calculated on Hellinger transformed data")  + 
  stat_boxplot(geom ='errorbar', lwd=1) + 
  theme(axis.title.y = element_text(size=12, face = "bold"), axis.text.y = element_text(size=16), axis.text.x = element_text(size=12, angle = 45, vjust = 0.95, hjust = 1), axis.title.x = element_blank(), panel.grid.major = element_blank(), panel.background = element_blank(), panel.border = element_blank(), axis.line = element_line(colour = "black"), title = element_text(size = 14, face = "bold"), legend.position = "none") +
ggtitle("Pairwise comparison of dissimilarities \nin paired and unpaired samples")

boxp
ggsave(boxp, device = "png", width = 30, height = 20, units = "cm", file ="/Users/kirstinebjergandersen/Desktop/boxp.png")

```

**Statistical test**
```{r}
wilcox.test(x=boxplotting$value[boxplotting$name=="Paired samples"],y=boxplotting$value[boxplotting$name=="Unpaired samples"],paried=FALSE)
```
**Test of median, min and max**
```{r}
boxplotting %>% group_by(name) %>%
  summarise(median=median(value),min=min(value),max=max(value))
```



